Three Problems for Sophistication

### 2015-01-16 Facticity/Facticity/paper.tex
...

@@

\begin{abstract}
Kolmogorov complexity is a sound measure of the amount information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise. Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed. This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been completely satisfactorily resolved.

@@

Before we start, we will define a framework into which we can fit all current variants of sophistication. The principle is always the same: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.

Let $\B = {0,1}^*$. We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}. Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}. A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set. A function $f$ is \emph{total} if $D_f = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix. \footnotemark

\footnotetext{Completeness is analogous to totality, in that a prefix-free Turing machine that does not halt corresponds to a complete prefix-free function.}

\begin{definition}[Models and model classes]
  A \emph{model} is a function $\phi:\B\to\B$. A
  \emph{model class} is a set of models. We distinguish four important
  model classes:
  \begin{itemize}
  \item $\C$ is the set of all partial computable functions,
  \item $\K\subset\C$ is the set of all prefix-free
    functions,
  \item $\T\subset\C$ is the set of all total functions,
  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
  \item $\F\subset\P$ contains, for every finite set $S$, a computable
    surjective function mapping the binary sequences of length
    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets, with uniform distributions.
  \end{itemize}
\end{definition}
\begin{definition}[Numberings]
A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$. We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$. We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
\end{definition}

@@

\begin{definition}[Complexity]
Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$. Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\} = \K$. Now define Kolmogorov complexity by
\begin{align*}
C(x)&=\min\{|\bar\imath| y:\phi_i(y)=x\},\\
K(x)&=\min\{|\bar\imath| y:\phi_i(y)=x, i\in\Pi\}.\\
\end{align*}
\end{definition}
This definition of $K$ is uncommon. It is functionally equivalent to the standard definition as used in \cite{li1993introduction}, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).

We now have the basic tools with which we set up our generic sophistication function. We first define the space of two-part representations for a given string $x$, and the codelength that each representation achieves:

@@

Define the following code-length functions for the representations:
\begin{itemize}
\item $L(m)$ is the length of a self-delimiting description of
  the model $m$, i.e. it is the length function of a prefix code for models. Each variant of sophistication must supply such a function.
\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
  required to describe $x$ using the best model in the model class.
\end{itemize}
\end{definition}
We can now define which of these representations express the data most efficiently:
\begin{definition}[Candidate set]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
\[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
\]
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$.
\end{definition}

\begin{definition}[Sophistication]
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}
For a function like sophistication to be meaningful, it must satisfy certain properties. We will check all currently published variants against the following minimal requirements:

\begin{description}
  \item[Invariance] Like the Kolmogorov complexity, the construction of a sophistication function makes certain ad-hoc choices, which are captured in the choice of an arbitrary universal Turing machine $U$ or an arbitrary numbering of $\C$. If we change what choice we make, the Kolmogorov complexity remains invariant up to a constant. This is the sole reason we allow ourselves to interpret the Komogorov complexioty as a property of the data, rather than simply the result of some computation on it. A similar invariance must hold for any sophistication function.
   \item[Unboundedness] Both the sophistication and the residual information (roughly the complexity minus the sophistication) must not be bounded by a constant. If the sophistication is bekow a constant for all strings it cannot have much meaning. Also, if it is always withing a constant of the Kolmogorov complexity, it cannot correspond to the notion we hope to capture with it.
   \item[Efficiency] The two-part representation $(m, y)$ that witnesses the sophistication (that is the one that reaches the minimum described by $\s^\M_\tau$) must reach the Kolmogorov complexity (or at least the difference should be insignificant). If there is an effective one-part representation that is much more efficient than the two-part representation that witnesses the sophistication, we cannot, in all honesty, claim that $(m, y)$ captures all the information in $x$.
\end{description}

Given the philosophical underpinnings of sophistication, these three requirements are the very least that we could expect. We will show that no current proposal satisfies all three.

For all variants but one, we show that all measures that are invariant are bounded. This leads us to the following conjecture. Let $\gamma(n)$ be a function representing a \emph{granularity}: the extent to which we care about differences between functions. That is, we will call two functions $f$ and $g$ $\gamma$-equal if $|f(x) - g(x)| \in O\left(\gamma\left(f(x)\right)\right)$.

\begin{conj}
Let $\s_g$ be a sophistication function constructed as in Definition~\ref{definition:sophistication}, built on some acceptable numbering $g$, and some fixed model class $\M$ and threshold function $\tau$ (omitted from the sub/superscript for the sake of clarity).

For any sophistication function so constructed, and any granularity $\gamma$ we have that for any two acceptable numberings $g$ and $h$, either:
\begin{itemize}
  \item there exist $\s_g$ and $\s_h$ which are not are $\gamma$-equal (ie $\s$ is not invariant),
  \item $s_g$ is $\gamma$-equal to $0$ or to the Kolmogorov complexity.
\end{itemize}
\end{conj}

In these terms, all instantiations of the Kolmogorov complexity, using different Universal Turing machines are $1$-equal. It may be unreasonable to require the sophistication to attain the same level of invariance (a log-invariant sophistication could still be useful). This conjecture allows such relaxations: if it is true, sophistication must be considered fundamentally flawed. We hope that this conjecture will prompt authors to address the matter of invariance critically, and to consider the possibility that sophistication and its ilk are an untenable proposition.

In the rest of this paper, we will analyse the existing variants one by one. We will conclude with the informal reasoning supporting this conjecture and consider its the consequences. What does this mean for the problem of meaningful information, what does it mean for algorithmic statistics and most importantly, what are the consequences for the idea of model selection in general?

@@

We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input. Can we assume that such a description, that is as efficient as the Kolmogorov complexity, creates a good separation into the structural part of the data and the noise?

The first problem that arises is that allowing any acceptable numbering can lead to inefficient model representations.

@@

No such function can be $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.

@@

In this proof, we deliberately use ridiculously inefficient index functions. Obviously one could simply pick a more efficient index; the point is that many definitions of sophistication to not mention any such restriction. As we shall see in the next, any inefficiency in the index will lead to a lack of invariance in sophistication.

@@

Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a UTM with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}). But this robustness property does \emph{not} carry over to definitions of sophistication, if we are interested in separating the model and its input, we must represent the model efficiently.

Another perspective ion this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content. In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\\
    K_\phi(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}. We will show that these values are invariant up to a constant (ie. complexity is an objective property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ to achieve the same inequality the other way around, completing the proof.
\end{proof}
This lemma shows that the complexity of a function depends on the choice of
enumeration by no more than a constant term.

There are two ways to use this notion of model complexity for more robust attempts to define sophistication, that differ in the way the two-part code is conceptualized. Both approaches are used in the literature, which is therefore sometimes confusing.
\begin{enumerate}
  \item On the one hand one can stick with Koppel's approach, but repair it by imposing the additional requirement that the used length of the indices used is equal to $C(\phi_i)$, and that the numbering is still acceptable. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}. In this case, the total code length is simply $C(x)$, and the sophistication is the length of the model index within this code. This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the model, which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$. The difference is that in this approach, we can construct everything on the basis of a single, unconstrained, acceptable numbering, without using additional contraints. It is also more natural in this setting to limit the model class of allowed functions. This approach is used for Kolmogorov's structure function \cite{cover1985kolmogorov}, Algorithmic Statistics \cite{gacs2001algorithmic}, Vi\'anyi's sophistication \cite{vitanyi2004meaningful} and Effective complexity \cite{gellmann1996information}.
\end{enumerate}

To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists a numbering for which the length of the index of a model is equal to its complexity, and that such a numbering is acceptable.

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}
This definition is a refinement of the one used in \cite{adriaans2012facticity}. That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable. Since we use numberings, we only need to show that they are acceptable, ie. we only need to be able to unpakc the efficient description, so the incomputability of the index function is no issue:
\begin{lemma}
  There is a faithful acceptable numbering.
\end{lemma}
\begin{proof}
Let $i_\tn{div}(y)$ be an index such that $\phi_{i_\tn{div}}(y)=\infty$ for all $y$. Define
  \[\psi_q=\begin{cases}
    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
  \]
  To show that $\psi$ is faithful, pick any function $f$. Then
\[\begin{split}
C(f)&=\min\{C(i):\phi_i=f\}\\
&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
\end{split}\]
This shows there is a sufficiently small $\psi$ index.

@@

\begin{lemma}
Let $L(\phi) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.

If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $s^\C$ is bounded for all numberings $\phi$
\end{lemma}
\begin{proof}
Let $\bar\imath p$ be any two-part representation for the data $x$, i.e. $\phi_i(p)=x$. Then construct an alternative two-part representation $\bar vi^* p$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* p)=\phi_{\psi_u(i^*)}(p) = \phi_i(p)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* p| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$. From this $i$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the facticity. Consequently, the facticity is bounded by $F^\phi(x)<|\overline{\imath_0}|$ for all $x$. 
\end{proof}
This result applies to Adriaans' facticity\cite{adriaans2012facticity}. In order to circumvent the problem, the UTM should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.

Note that the condition for this lemma applies to all known partial computable prefix functions. We conjecture that it applies to all:

\begin{conjecture}
For any partial computable prefix code length function $L$, the function
\[
\min\{L(i)-K(i):i\ge i_0\}
\]
is unbounded in $i_0$.
\end{conjecture}

This result shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}). In the cases where the data representations are taken as a single code word, the model-part must represent the model as efficient as $K(f)$ does. Instead of a numbering, we require a self-delimiting code for functions rather than a numbering.

In the rest of this paper, we will take the second perspective: we will consider the representations as two-part descriptions $K(f)$, and $|y|$, such that $f(y)=x$. This description can be unpacked by a very specific universal Turing machine, but contrary to the construction of the Kolmogorov complexity, we are not free to choose any UTM.

@@

On the other end of the spectrum, we find the problem of optimally performing, highly-specific models. Most poigniantly, the singletons models. Sets which only contain $x$, probability distributions with $p(x) = 1$, or functions which produce $x$ for all arguments. Lemma~\ref{lemma:thecoolone} allows us to give these models a constant boost as well:

\begin{lemma}[overfitting]
Let $\M$ be a prefix model class where for every $x\in\X$ there is a singleton model $f\in\M$ with $f(\epsilon)=x$. Then there is an enumeration $\phi_1,\phi_2,\ldots$ of the prefix partial recursive functions, and a constant $c$, such that
\[
K(x)-S^{\M,\phi}(x)\le c
\]
for all $x\in\X$.
\end{lemma}
\begin{proof}
Let $\psi_1,\psi_2,\ldots$ be any default enumeration of the partial recursive prefix functions. Note that since $f$ is a prefix function, if $f$ is defined for input $\epsilon$ then it cannot be defined for any other input. Pick $f,x$ with $f(\epsilon)=x$. Note that $x$ can be computed from $f$ and a fixed program, so there is a $c$ such that $K(x)\le K(f)+c$. Vice versa, given any $x$ we can construct an index of $f$, since $\psi$ is an acceptable numbering. Therefore $|K(f)-K(x)|\le c$.

We now define a computable function $D$ by $D(\bar\imath p)=j$ where $\psi_j(\epsilon) = \psi_i(p)$.  We will show that the two conditions of Lemma~\ref{lem:thecoolone} hold for the prefix function $D$.

Then by Lemma~\ref{lem:thecoolone} there is an enumeration $\phi_1,\phi_2,\ldots$ such that $S^{\M,\phi}(x)=|\bar 0|+S^{\M',\psi}(x)$. We observed that $|K(f)-K(x)|\le c$ for any $f\in\M'$, so $S^{\M',\psi}(x)\ge K(x)+c$. This proves the lemma.
\end{proof}

Many variants in the literature are susceptible to this problem, including the structure function \cite{cover1985kolmogorov,gacs2001algorithmic}. \cite{vitanyi2004meaningful} and \cite{adriaans2012facticity} circumvent this problem by using non-self-delimiting representations. Since a singleton representation places all information in the (self-delimiting) model part, singletons are automatically less efficient than representations placing some information in the input. While this side-steps the issue of the singletons, there is no suggestion that this makes the sophistication invariant.

Another solution would be to make the constant in the threshold function dependent on the choice of enumeration. We are not aware of any approach to this effect, and there does not seem to be a simple way to determine the thresholkd based on the properties of the numbering.

In the previous section, we saw a first example of \emph{underfitting}: a second UTM with a more efficient model representation became a model for all strings. Even if we chose our model-description carefully, to avoid the issues of the previous section, however, universal Turing machines are still candidate model for the data. These candidates represent the data efficiently (they reach the Kolmogorov complexity up to a constant), but they are extremely generic: taking a second sample from the model can produce a string that has nothing to do with the data. Nevertheless, these models perform well, in terms of compression, and there is no justification for excluding them from the candidate set.

The opposite problem is \emph{overfitting}. A model that can produce only $x$ can be described in as many bits as it takes to describe $x$, plus some constant amount. Again, such models perform optimally, but they are not very informative. We call this \emph{overfitting}.

As we shall see, most variants of sophistication are constructed specifically to prevent the problems of over- and underfitting.

Any sophistication function built on a model class that includes singletons, or universal models is in danger of over or underfitting. In that case, we can construct a numbering, so that a particular model, or class of models, represents all data more efficiently than any other model by an arbitrary constant. This constant boost is enough to force all other models out of the candidate set.

The following lemma generalizes the principle. We will instantiate it for different variants of sophistication:

@@

\section{=======Stuff to move up or delete=======}

\section{Old structure}

\subsection{Koppel and Atlan}

>>>>

% --- DELETE BEFORE SUBMUSSIONS ---
\pagestyle{headings}

@@

\begin{abstract}
Kolmogorov complexity is a sound measure of the amount of information in a binary string, but it does not distinguish between the information that expresses the structural properties of the data, and noise. Kolmogorov's definition of the \emph{structure function} in his lecture Tallinn in 1973 was the first attempt to separate meaningful information from noise, but many alternative definitions of the \emph{sophistication}, or the amount of meaningful information in the data, have subsequently been proposed. This indicates that the idea is considered worth pursuing by multiple distinguished scholars, but at the same time it shows that the issue has never been completely satisfactorily resolved.

@@

We will generalizes the main proposals for sophistication into a single, consistent notation. The unifying principle is this: we start with a class of models, we consider the space of two-part descriptions using these models, we select the set of representations with the shortest code-lengths and we select, by some method, the best model from this set.

Let $\B = \{0,1\}^*$. We deal with partial functions $\phi: \B \to \B$, which we will also call \emph{models}. Such functions are \emph{computable} if they can be computed by a Turing machine as described in \cite[Definition~1.7.1]{li1993introduction}. A function is prefix-free if its domain, $D_f = \{y : f(y) \neq \infty\}$ is a prefix set. A function $f$ is \emph{total} if $D_f = \B$. A prefix-free function is \emph{complete} if every infinite binary string has a member of $D_f$ as a prefix. \footnotemark

\footnotetext{Completeness is analogous to totality: a total function corresponds to a Turing machine that halts for all its inputs, while a complete function corresponds to a Turing machine that always halts, so long as there are enough bits on the input tape.}

As noted in the introduction, the Kolmogorov complexity is invariant to the choice of description language. Formally, our description language creates a numbering of the partial computable functions.

\begin{definition}[Numberings]
A numbering is an enumeration of $\C$, denoted as $\phi_1, \phi_2, \ldots$ or by a function $g: \N \to \C$. We fix one canonical numbering $g^\circ$, which works as an effective description, that is for any $i \in \N, y\in \B$, we can effectively compute $(g^\circ(i))(y)$. We call a numbering $g$ \emph{acceptable} if there exist total, computable functions $a, b: \N \to \N$ as that for all $i$ $g^\circ(a(i)) = h(i)$ and  $g^\circ(i) = h(b(i))$.
\end{definition}

When we say that Kolmogorov complexity is invariant, we mean that changing the canonical acceptable numbering changes the Kolmogorov complexity by at most a constant.

@@

\begin{definition}[Complexity]
Let $\phi_1,\phi_2,\ldots$ be the canonical acceptable enumeration of $\C$. Let $\Pi$ be a computable subset of the natural numbers such that  $\{\phi_i:i\in\Pi\}$ corresponfs to the set of all computable prefix functions. Now define Kolmogorov complexity by
\begin{align*}
C(x)&=\min\{|\bar\imath y|:\phi_i(y)=x\},\\
K(x)&=\min\{|\bar\imath y|:\phi_i(y)=x, i\in\Pi\}.\\
\end{align*}
\end{definition}
This definition of $K$ is uncommon. It is equivalent to the standard definition as used in \cite{li1993introduction}, up to a constant, but this form is more convenient for defining the $K$-complexity of non-prefix functions (see Definition~\ref{definition:model-complexity} below).

We now have the basic tools with which we set up our generic sophistication function. We start with the class of allowed models:

\begin{definition}[Models and model classes] A \emph{model class} is a set of models. We distinguish four important model classes:
  \begin{itemize}
  \item $\C$ is the set of all partial computable functions,
  \item $\K\subset\C$ is the set of all prefix-free
    functions,
  \item $\T\subset\C$ is the set of all total functions,
  \item $\P\subset\K$ is the set of all complete, prefix-free functions,
  \item $\F\subset\P$ contains, for every finite set $S$, a computable
    surjective function mapping the binary sequences of length
    $\lceil\log S\rceil$ onto $S$. In other words, this is the model class of finite sets, with uniform distributions.
  \end{itemize}
\end{definition}

We note that the model classes $\T$ and $\P$ are unusual in that they are not enumerable.

@@

Define the following code-length functions for the representations:
\begin{itemize}
\item $L(m)$ is the length of a self-delimiting description of
  the model $m$, i.e. it is the length function of a prefix code for models. Each variant of sophistication must supply such a function.
\item $L^m(x)=L(m)+\min\{|y|:m(y)=x\}$ measures the number of bits required to describe both the model $m$ and the data $x$ using the model.
\item $L^\M(x)=\min\{L^m(x):m\in\M\}$ measures the number of bits
  required to describe $x$ using the best model in the model class.
\end{itemize}
\end{definition}
We can now define which of these representations express the data most efficiently:
\begin{definition}[Candidate set]
The \emph{candidate models} for $x$ from model class $\M$ and threshold function $\tau$ are
\[
  C^\M_\tau(x):=\{m\in\M:L^m(x)\le \tau(x) + c\}
\]
with $c$ a constant independent of $x$ which may be chosen arbitrarily. The default threshold function is $\tau^\circ(x)=L^\M(x)$. We will often omit the subscript if the default threshold function is used. 
\end{definition}

\begin{definition}[Sophistication]
  The model complexity, meaningful information, effective complexity, facticity or \emph{sophistication},  of a string $x$ given a model class $\M$ and threshold function $\tau$ is
  \[
  \s_\tau^{\M}(x):=\min\{L(m):m\in C^\M_\tau(x)\}\p
  \]\label{definition:sophistication}
\end{definition}

Finally, we must formalize what we mean when we say that we desire sophistication to be invariant. For the Kolmogorov complexity, we know that a change in acceptable numbering causes no more than a constant change. It may be that a less stringent invariance (for instance a $\log$ term) can still yield a worthwhile definition of sophistication. To allow such distinctions, we generalize the notion of invariance:

\begin{definition}
Let $\gamma:\N\to \N$ be some nondecreasing function. Two functions $f$ and $g$ are $\gamma$-equivalent if $|f(x) - g(x)| \in O(\gamma(\max(f(x), g(x)))$.
\end{definition}

We say that a formalization of sophistication $s_\phi$ built on acceptable numbering $\phi$ is $\gamma$-invariant to the choice of numbering is $s_\phi$ and $s_\psi$ are $\gamma$-equal for any two acceptable numberings $\phi$ and $\psi$. Thus, by this definition, Kolmogorov complexity is $1$-invariant.

@@

We start from a simple intuition: the definition of Kolmogorov complexity already uses a description of the data in two parts; a computable function and and input. Can we assume that when such a description is as short as the Kolmogorov complexity, it automatically creates a good separation into the structural part of the data and the noise? As we will show, the representation of the structural part plays a fundamental part in answering this question. If it is in any way inefficient, the resulting sophistication function does not behave as expected.

The issues can be very subtle and require a careful attention that is not always afforded in the literature. We will first illustrate what happens if any acceptable numbering is allowed:

@@

No such function can be both $\gamma$-invariant and unbounded for any $\gamma(n) \in o(n)$.
\end{lemma}
\begin{proof}
Choose some $\gamma(n) \in o(n)$. Assume towards contradiction that the sophistication is $\gamma$-invariant and unbounded.

@@

In this proof, we deliberately used ridiculously inefficient index functions. Obviously, one could simply pick a more efficient index; the point is that many definitions of sophistication do not mention any such restriction. As we shall see, any inefficiency in the index will lead to a lack of invariance in sophistication.

@@

Interestingly, using an inefficient index is no problem in the definition of Kolmogorov complexity, since we can select as a model a universal Turing machine with an efficient index at only a constant penalty (by the Invariance Theorem \cite{li1993introduction}). But this robustness property does \emph{not} carry over to definitions of sophistication. If we are interested in separating the model and its input, we must represent the model efficiently.

Another perspective on this problem is that the length of a model in an ordinary description does not represent its complexity, much like the length of a compressible string does not reflect its information content. In order to obtain a more robust measure of model complexity, we describe a natural extension of Kolmogorov complexity to functions, and show that this retains invariance.

\begin{definition}
  The complexity of a partial recursive function is defined by
  \begin{align*}
    C_\phi(f) &:= \min\{C(i):\phi_i=f\}\\
    K_\phi(f) &:= \min\{K(i):\phi_i=f\}.
  \end{align*} \label{definition:model-complexity}
\end{definition}
These definitions also appear in \cite{grunwald2004shannon} and \cite{vitanyi2004meaningful}. We will show that these values are invariant up to a constant (ie. complexity is an objective property of a function):
\begin{lemma}[Invariance]
Let $\phi_1, \phi_2, \ldots$ and $\psi_1, \psi_2,\ldots$ be two acceptable numberings. There exists a constant $c$ such that $\left| K^\phi(f) - K^\psi(f)\right | \leq c$ for all $f$. \label{lemma:invariance}
\end{lemma}
\begin{proof}
Let $g(i)$ be the function which which translates from $\psi$ to $\phi$.
\begin{align*}
K^\phi(f) &= \min\left\{ K^\phi(i) : \phi_i= f\right\} \\
&\geq \min\left\{ K^\psi(i) : \phi_i= f\right\} - c\\
&= \min\left\{ K^\psi(i) : \psi_{g(i)}= f\right\} - c\\
&\geq \min\left\{ K^\psi(i) : \psi_i= f\right\} - c\\
&= K^\psi(f).
\end{align*}
We can reverse $\phi$ and $\psi$ to achieve the same inequality the other way around, completing the proof.
\end{proof}
This lemma shows that the complexity of a function does not depend on the choice of
enumeration by more than a constant term.

There are two ways to use this notion of model complexity for more robust attempts to define sophistication. Confusingly, both approaches are used in the literature,
\begin{enumerate}
  \item One can stick with the approach of using the model used by the representation that witnesses the Kolmogorov complexity, but repair it by imposing the additional constraint that an efficient numbering is used. Specifically, that the length of the indices satisfies $|i| = C(\phi_i)$, and that the numbering is still acceptable. We construct such an index below. This approach is taken by Adriaans \cite{adriaans2012facticity}. In this case, the total code length is simply $C(x)$, and the sophistication is the length of the model index within this code. This approach is the most direct formalization of the intuition that the two part code used within the definition of Kolmogorov complexity separates structure from noise.
  \item On the other hand, rather than distinguishing the two components of the code \emph{within} the definition of Kolmogorov complexity, we can also use Kolmogorov complexity as a black box to measure the complexity of the \emph{model}, which is then the first part of a two-part code describing the data. With this approach, the total code length is $K(f)+|y|$, where $y$ is an input to the model $f$ such that $f(y)=x$. The difference is that in this approach, we can construct everything on the basis of a single acceptable numbering, without using additional contraints. It is also more natural in this setting to limit the model class of allowed functions. This approach is used for Kolmogorov's structure function \cite{cover1985kolmogorov}, Algorithmic Statistics \cite{gacs2001algorithmic}, Vi\'anyi's sophistication \cite{vitanyi2004meaningful} and Effective complexity \cite{gellmann1996information}.
\end{enumerate}

To show that the two approaches are equivalent with the correct restriction on the index, we must show that there exists an acceptable numbering for which the length of the index of a model is equal to its complexity.

\begin{definition}[Faithful Numbering]\label{def:faithful}
  A numbering $\psi_1,\psi_2,\ldots$ of the partial recursive
  functions is \emph{faithful} if there is a constant $c$ such that
  for all indices $i$ there is a $j$ such that $\psi_i=\psi_j$ and
  $|j|\le C(\psi_j)+c$.
\end{definition}
This definition is a refinement of the one used in \cite{adriaans2012facticity}. That publication defines an \emph{index function}, rather than a numbering, and notes that it is not computable. Since a numbering is all that is needed, we only need to show acceptability. The incomputability of the index function is no issue:
\begin{lemma}
  There is a faithful acceptable numbering.
\end{lemma}
\begin{proof}
Let $\tn{div} \in \N$ be an index such that $\phi_{\tn{div}}(y)=\infty$ for all $y$. Define
  \[\psi_q=\begin{cases}
    \phi_{\phi_i(p)}&\tn{if $q$ can be written as $\bar\imath p$ and $\phi_i(p)<\infty$,}\\
    \phi_{i_\tn{div}}&\tn{otherwise.}\end{cases}
  \]
  To show that $\psi$ is faithful, pick any function $f$. Then
\[\begin{split}
C(f)&=\min\{C(i):\phi_i=f\}\\
&=\min\{\min\{|\bar a b|:\phi_a(b)=i\}:\phi_i=f\}\\
&=\min\{|\bar a b|:\phi_{\phi_a(b)}=f\}\\
&=\min\{|\bar a b|:\psi_{\bar a b}=f\}.
\end{split}\]
This shows there is a sufficiently small $\psi$ index.

@@

\begin{lemma}
Let $L(\phi_i) = |\bar\imath|$ for a faithful index, and some prefix encoding function, and let $\s^\C$ be defined on this $L$ with the default threshold function.

If $\min\{i\ge i_0:|\bar\imath|-K(i)\}$ is an unbounded function of $i_0$, then $\s^\C$ is bounded for all numberings $\phi$.\label{lemma:prefix-inefficiency}
\end{lemma}
\begin{proof}
Let $\bar\imath y$ be any two-part representation for the data $x$, i.e. $\phi_i(y)=x$. Then construct an alternative two-part representation $\bar vi^* y$, where $i^*$ is the shortest $\psi_u$-program for $i$ such that $\phi_v(i^* y)=\phi_{\psi_u(i^*)}(y) = \phi_i(y)=x$. We compare the lengths of these two representations. Note that $|i^*|=K^\psi(i)$. Therefore,
\[
|\bar\imath p|-|\bar v i^* y| = |\bar\imath|-|\bar v| - |i^*| = |\bar\imath|-|\bar v|-K(i).
\]
By assumption there must be an $i_0$ such that the above expression is positive for all $i>i_0$. From this $i_0$ onwards, the second representation (using $v$) will have a shorter code length, so $\bar\imath p$ cannot achieve the minimum in the definition of the sophistication. Consequently, the sophistication is bounded by $\s(x)<|\overline{\imath_0}|$ for all $x$. 
\end{proof}
This result applies to Adriaans' facticity\cite{adriaans2012facticity}. In order to circumvent the problem, the UTM should should encode the model $f$ in $K(f)$ bits, relying on $K$'s own prefix encoding to delimit the model from its input.

Note that the condition for this lemma applies to all commonly used prefix encoding functions. We conjecture that it applies to all possible computable prefix encoding functions:

\begin{conjecture}
For any partial computable prefix codelength function $L$, the function
\[
\min\{L(i)-K(i):i\ge i_0\}
\]
is unbounded in $i_0$.
\end{conjecture}

Lemma~\ref{lemma:prefix-inefficiency} shows that even if we have a faithful numbering, we can only use it in settings where we don't count the cost of delimiting the model from the input (such as \cite{koppelSoph1988}). In the cases where the data representations are taken as a single code word, the model-part, including delimiting information, must represent the model as efficiently as $K(f)$ does.

In the rest of this paper, we will avoid these issues and take the second perspective: we will consider the representations as two-part descriptions $(f, y)$ with codelength $K(f) + |y|$, such that $f(y)=x$.

@@

We now come to the main issue plaguing sophistication. In the naive setting, where we allow all prefix-free functions as models, and use the default threshold function, the candidate set always contains two representations that clearly do not provide a good separation into structure and content. The first is a universal model: a universal Turing machine which simply runs its shortest program for $x$. The second is a \emph{singleton model} a model which returns $x$ for any input. In the first representation all information about $x$ is stored in the second part of the two-part description. In the second representation all information is stored in the model. The model complexity suggested by these representations is either constant, or equal to the Kolmogorov complexity.

The key problem is that both can be present in the candidate set, and the formalization of sophistication must ensure that neither are selected. As we will show, in many settings, we can ensure that they are the \emph{only} models in the candidate set. The principle we use is that the threshold on the candidate set is determined by a constant, and a change in numbering affects model size by a constant. Thus, We can choose our numbering  so that all but a certain class of models take a constant penalty, pushing them out of the candidate set. Since we use this trick repeatedly we have generalized it into an (admittedly complex) lemma.

@@
